Near and Mid-infrared properties of known $z\geq5$ Quasars

### 2019-06-11 tex/VHzQ.tex
tex/VHzQ.tex

There are 14 objects that have no NIR detections.  3 of these
(PSOJ053.9605-15.7956, PSO J056.7168-16.4769 and DELSJ0411-0907) have
been observed (by VHS) but are out of our queried time range 
(for which the data is publicly available). 6 objects have not been 
observed (or at least the data is not in the
WSA/VSA archives yet) and 5 objects are at a declination $\delta \geq +60$
deg are too far north for UKIRT and cannot be observed.

  \subsubsection{Comparing WFCAM and VISTA}
  There are 130 overlapping QSOs between WFCAM and VISTA. We used the
  {\tt VegaToAB} values to put these objects on the same AB system, and
  for each object compared the two measurements. First, we calculated the
  weighted average (calibrated flux) in each filter of both and
  calculated the ratio and difference between each measurement and the
  average.  Then for each filter we calculated the weighted average of
  the differences (in mag) for each instrument to see if there were
  significant offsets. The results are given in
  Table~\ref{tab:WFCAM_vs_VISTA}.  The only filter with a significant
  offset is the $Y$-band. All of the VISTA averages are negative and all
  of the WFCAM ones are positive. The $Ks$ versus $K$ band comparison may be
  affected the different shapes of the filters, with $K$ being significantly 
  wider than $K_s$, able to detect light at longer wavelengths.
  \begin{table}
    \centering
    \begin{tabular}{l r r}
      \hline  \hline
      abs(VIRCAM & \multirow{2}{*}{millimags} &  no. of  \\
      -  WFCAM)      &                                        &  objects \\
      \hline
      $Z$                 &  23.2 	& 3 \\
      $Y$                 &  57.3 	& 53 \\
      $J$                  &    2.1 	& 106 \\
      $H$                 &  45.8     &  96 \\
      $K_{\rm s}$/$K$ &  25.2     & 110 \\
      \hline  \hline
    \end{tabular}
    \caption{Comparing the magnitudes in different WFCAM/UKIRT and 
      VIRCAM/VISTA near-infrared bands.}
    \label{tab:WFCAM_vs_VISTA}
  \end{table}

@@

\begin{figure}
  %% trim=l b r t
  \includegraphics[width=8.6cm, clip,trim=6mm 6mm 0mm 6mm]
  {/cos_pc19a_npr/programs/quasars/highest_z/detections/WISEmag_vs_coverage_20190605.pdf}
  \centering
  \vspace{-14pt}
  \caption[]{WISE W1/2/3/4 {\tt w$x$mpro} magnitude against signal-to-noise,
    colour coded by {\tt w$x$cov} the mean coverage depth, in each corresponding band.}
  \label{fig:WISEmag_vs_coverage}
\end{figure}
Before reporting on the detection rates, we investigate this effect
using the AllWISE Source Catalog. Figure~\ref{fig:WISEmag_vs_coverage}
shows the WISE ALLWISE magnitude versus signal-to-noise, colour coded
by {\tt w$x$cov} the mean coverage depth, in each corresponding
band. In the two shorter bands W1/2 we see the clear and expected
trend for brighter objects to have larger SNR, and also for the higher
signal to noise for objects with more exposures at a given
magnitude. The behaviour for the W3/4 bands is different, with two
populations clearly evident in W3 and although a bit more mixed, also
in W4. With the suggested split at SNR$>2$, and no obvious
R.A./Declination dependence seen, this behaviour is explained by the
fact that there are non-detections in W3/4 for objects (with high W1/2
SNR) that are reported in the ALLWISE catalogue.

@@

\begin{figure}
  %% trim=l b r t
  \includegraphics[width=8.6cm, clip, trim=2mm 0mm 2mm 0mm]
  {/cos_pc19a_npr/programs/quasars/highest_z/detections/WISEsnrW1W2W3W4_2by3_v1.pdf}
  \centering
  \vspace{-14pt}
  \caption[]{WISE signal-to-noise measures for the four bands, as well
    as for (W1-W2) colour.  The points are colour coded by redshift.}
  \label{fig:WISEmag_vs_coverage}
\end{figure}

Recently, \citet{Assef2018} released two large catalogues of AGN
candidates identified across 30,000 deg$^2$ of extragalactic sky from
the WISE AllWISE Data Release. The ``R90'' catalogue, contains 4.5M
AGN candidates at 90\% reliability (and $\approx$150 AGN candidates
per deg$^2$) while the ``C75'' catalogue consists of 20.9M AGN
candidates at 75\% completeness (and ($\approx$700 AGN candidates per
deg$^2$). Crossmatching our catalogue of 463 VH$z$Qs with these
catalogues, produces 42 matches with the R90 sample and 98 matches
with the C75 sample. Both catalogues unsurprisingly match to the
ultraluminous quasar SDSS J0100+2802 \citep{Wu2015} while the C75, but
not the R90 catalogue matches to ULAS J1120+0641
\citep{Mortlock2011}. Neither catalogue matches to J1342+0928
\citep{Banados2018}.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/UHS_J0439+1634.png} 
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J1443+3623.png}
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J1623+4705.png}
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J0100+2802.png}
    \caption{Four of the brightest MIR VH$z$Qs. 
      The thumbnails are the RGB colour outputs using the ALLWISE W1W2W3 bands from the 
      IRSA WISE Image Service, with a scale of 120'' on the side. 
      {\it (a)} UHS J0439+1634 was discovered by
      \citet{Fan2019} and is a strongly lensed quasar at $z=$6.51. 
      {\it (b)} SDSS J0100+2802 is the \citet{Wu2015} object. 
      {\it (c)} SDSS J1443+3623 has {\tt w3mpro, w4mpro } = ($11.189\pm0.1$, $8.699\pm0.26$) and 
      {\it (d)} SDSS J1623+4705 has {\tt w3mpro, w4mpro } = ($11.425\pm0.1$, $8.954\pm0.28$).}
    \label{fig:VHzQ_W4s}
\end{figure*}
The set of VH$z$Q detected in WISE W3 and W4 contains 51 objects that are
detected in the broad W3 filter and 10 objects that are formally detected
in W4 ({\tt w4snr}$\geq$3.0). Four of the brightest W4 objects are
presented in \label{fig:VHzQ_W4s}. UHS J0439+1634
(Fig.~\ref{fig:VHzQ_W4s} (a)) with W4=7.165$\pm$0.12 was discovered by
\citet{Fan2019} and is a strongly lensed quasar at $z=$6.51. This high
luminosity is mostly not intrinsic, but is boosted by an intervening
redshift $z\sim$0.7 galaxy. SDSS J0100+2802 (redshift $z=6.33$) does
not have a formal W4 detection (W4 $=8.98\pm0.45$ and {\tt w4snr}=2.4),
but as reported by \citet{Wu2015}, has a bright detection in the
2MASS $J$, $H$ and $K_{\rm s}$ bands and we report it here.

>>>>

There are 14 objects that have no NIR detections. 3 of these
(PSOJ053.9605-15.7956, PSO J056.7168-16.4769 and DELSJ0411-0907) have
been observed (by VHS) but are out of our queried time range (for
which the data is publicly available). 6 objects are at declination
$\delta<0$ deg and have not been observed (or at least the data is not
in the VSA archive yet). 5 objects are at a declination $\delta \geq
+60$ deg are too far north for UKIRT and cannot be observed.

  \subsubsection{Comparing WFCAM and VISTA}
  There are 130 quasars observed in the overlapping area between WFCAM
  and VISTA. We used the {\tt VegaToAB} values to put these objects on
  the same AB system, and for each object compared the two
  measurements. First, we calculated the weighted average (calibrated
  flux) in each filter of both and calculated the ratio and difference
  between each measurement and the average.  Then for each filter we
  calculated the weighted average of the differences (in mag) for each
  instrument to see if there were significant offsets. The results are
  given in Table~\ref{tab:WFCAM_vs_VISTA}.  The only filter with a
  significant offset is the $Y$-band. All of the VISTA averages are
  negative and all of the WFCAM ones are positive. The $Ks$ versus $K$
  band comparison may be affected the different shapes of the filters,
  with $K$ being significantly wider than $K_s$, able to detect light at
  longer wavelengths.
  \begin{table}
    \centering
    \begin{tabular}{l r r}
      \hline  \hline
      abs(VIRCAM & \multirow{2}{*}{millimags} &  no. of  \\
      -  WFCAM)      &                                        &  objects \\
      \hline
      $Z$                 &  23.2 	& 3 \\
      $Y$                 &  57.3 	& 53 \\
      $J$                  &    2.1 	& 106 \\
      $H$                 &  45.8     &  96 \\
      $K_{\rm s}$/$K$ &  25.2     & 110 \\
      \hline  \hline
    \end{tabular}
    \caption{Comparing the magnitudes in different WFCAM/UKIRT and 
      VIRCAM/VISTA near-infrared bands.}
    \label{tab:WFCAM_vs_VISTA}
  \end{table}

@@

\begin{figure}
  %% trim=l b r t
  \includegraphics[width=8.6cm, clip,trim=6mm 6mm 0mm 6mm]
  {/cos_pc19a_npr/programs/quasars/highest_z/detections/WISEmag_vs_coverage_20190605.pdf}
  \centering
  \vspace{-14pt}
  \caption[]{WISE W1/2/3/4 {\tt w$x$mpro} magnitude against signal-to-noise,
    colour coded by {\tt w$x$cov} the mean coverage depth, in each corresponding band.}
  \label{fig:WISEmag_vs_coverage}
\end{figure}
Before reporting on the detection rates, we investigate this effect using the AllWISE Source Catalog. Figure~\ref{fig:WISEmag_vs_coverage} shows the WISE ALLWISE magnitude versus signal-to-noise, colour coded by {\tt w$x$cov} the mean coverage depth, in each corresponding band. The {\tt w$123$cov} values are the mean pixel coverage in W1/2/3 from the W1/2/3 Atlas Tile Coverage Map within an aperture of circular area with a radius of 8.25'' centered on the position of this source. For {\tt w$4$cov} this radius is 16.5'' (the AllWISE Source Catalog \href{http://wise2.ipac.caltech.edu/docs/release/allwise/expsup/sec2_1a.html#w2cov}{Column Descriptions} has further details). The {\tt w$x$cov} value takes into account e.g., individual pixels in the measurement area that may be masked or otherwise unusable (reducing the effective pixel count and thus the mean coverage value) as well as pixels that are affected by distortions across the across the focal plane in single-exposure images (where this distortion is corrected when coadding to generate the Atlas Images).

In the two shorter bands W1/2 we see the clear and expected
trend for brighter objects to have larger SNR, and also for the higher
signal to noise for objects with more exposures at a given
magnitude. The behaviour for the W3/4 bands is different, with two
populations clearly evident in W3 and although a bit more mixed, also
in W4. With the suggested split at SNR$>2$, and no obvious
R.A./Declination dependence seen, this behaviour is explained by the
fact that there are non-detections in W3/4 for objects (with high W1/2
SNR) that are reported in the ALLWISE catalogue.

@@

\begin{figure}
  %% trim=l b r t
  \includegraphics[width=8.6cm, clip, trim=2mm 0mm 2mm 0mm]
  {/cos_pc19a_npr/programs/quasars/highest_z/detections/WISEsnrW1W2W3W4_2by3_v1.pdf}
  \centering
  \vspace{-14pt}
  \caption[]{WISE signal-to-noise measures for the four bands, as well
    as for (W1-W2) colour.  The points are colour coded by redshift.}
  \label{fig:WISEmag_vs_snr}
\end{figure}

Recently, \citet{Assef2018} released two large catalogues of AGN
candidates identified across 30,000 deg$^2$ of extragalactic sky from
the WISE AllWISE Data Release. The ``R90'' catalogue, contains 4.5M
AGN candidates at 90\% reliability (and $\approx$150 AGN candidates
per deg$^2$) while the ``C75'' catalogue consists of 20.9M AGN
candidates at 75\% completeness (and ($\approx$700 AGN candidates per
deg$^2$). Crossmatching our catalogue of 463 VH$z$Qs with these
catalogues, produces 42 matches with the R90 sample and 98 matches
with the C75 sample. Both catalogues unsurprisingly match to the
ultraluminous quasar SDSS J0100+2802 \citep{Wu2015} while the C75, but
not the R90 catalogue matches to the first redshift $z>7$ quasar ULAS J1120+0641
\citep{Mortlock2011}. Neither catalogue matches to the current highest-redshift 
object  J1342+0928 \citep{Banados2018}.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/UHS_J0439+1634.png} 
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J1443+3623.png}
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J1623+4705.png}
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J0100+2802.png}
    \caption{Four of the brightest MIR VH$z$Qs. 
      The thumbnails are the RGB colour outputs using the ALLWISE W1W2W3 bands from the 
      IRSA WISE Image Service, with a scale of 120'' on the side. 
      {\it (a)} UHS J0439+1634 was discovered by
      \citet{Fan2019} and is a strongly lensed quasar at $z=$6.51. 
      {\it (b)} SDSS J0100+2802 is the \citet{Wu2015} object. 
      {\it (c)} SDSS J1443+3623 has {\tt w3mpro, w4mpro } = ($11.189\pm0.1$, $8.699\pm0.26$) and 
      {\it (d)} SDSS J1623+4705 has {\tt w3mpro, w4mpro } = ($11.425\pm0.1$, $8.954\pm0.28$).}
    \label{fig:VHzQ_W4s}
\end{figure*}
The set of VH$z$Q detected in WISE W3 and W4 contains 51 objects that are
detected in the broad W3 filter and 10 objects that are formally detected
in W4 ({\tt w4snr}$\geq$3.0). Four of the brightest W4 objects are
presented in \label{fig:VHzQ_W4s}. UHS J0439+1634
(Fig.~\ref{fig:VHzQ_W4s} (a)) with W4=7.165$\pm$0.12 was discovered by
\citet{Fan2019} and is a strongly lensed quasar at $z=$6.51. This high
luminosity is mostly not intrinsic, but is boosted by an intervening
redshift $z\sim$0.7 galaxy. SDSS J0100+2802 (redshift $z=6.33$) does
not have a formal W4 detection (W4 $=8.98\pm0.45$ and {\tt w4snr}=2.4),
but as reported by \citet{Wu2015}, has a bright detection in the
2MASS $J$, $H$ and $K_{\rm s}$ bands and we report it here.
%%
Other high-redshift quasars that are bright W3/4 objects may well 
also so be lensed, but high-resolution follow-up is needed to confirm.
